---
title: "data_exploration"
author: "Justin Sarna"
date: "April 14, 2017"
output: html_document
---
```{r}
# Load the necessary libraries
library(ggplot2)
library(useful)
library(caret)
library(ISLR)
library(scales)
library(plyr)
library(rpart)
library(rpart.plot)
library(class)
library(MuMIn)
library(caret)
```
# Create new variables for potential use in visualizations and model specification
```{r}
# code from hw4 to be used for creating additional models
acs$HighIncome <- as.numeric(with(acs, FamilyIncome >= 250000))
acs$foodstamp_binary <- ifelse(acs$FoodStamp == "Yes",1,0) # (yes = 1, no = 0)
acs$own_home <- ifelse(acs$OwnRent == "Rented",0, ifelse(acs$FamilyIncome == "Mortgage",1,2)) # (own = 1, rent = 0)
acs$family_type_cat <- ifelse(acs$FamilyType == "Married",1, ifelse(acs$FamilyIncome == "Female Head",2,3))
# married = 1, male head = 2, female head = 3
acs$InsuranceHigh <- (acs$Insurance > 1000) * 1
acs$NumWorkers2 <- (acs$NumWorkers == 2) * 1
acs$HouseCostsHigh <- (acs$HouseCosts > 1000) * 1
acs$high_electric <- (acs$ElectricBill > 350) * 1
# Break it into a training and test set with an 80/20 split.
set.seed(447)
testrecs <- sample(nrow(acs),0.2 * nrow(acs))
acs_test <- acs[testrecs,]
acs_fit <- acs[-testrecs,]  
# Create binary variable where 1 = not on food stamps & not renting & married
acs$HI_pred1 <- 0
acs$HI_pred1[acs_test$FoodStamp == 'No' & acs_test$OwnRent != 'Rented' & acs_test$FamilyType == 'Married'] <- 1
acs$null_model <- 0
```

```{r}
# I like this visualization for a quick visual of what I'm dealing with before digging in (modified from class notes)
ggplot(acs,aes(x=FamilyIncome)) + geom_density(fill="#31a354", color="#31a354") +
  geom_vline(xintercept=250000) + scale_x_continuous(label=multiple.dollar, limits=c(0,1000000))
```

### Interesting test I found for testing data normality - downside is it has a 5000 observation limit
If p-value is less than .05 (or chosen significance level) then sample is NOT normally distributed
This is not relevant here, but worth keeping for future
The plot shows that there is a clear left skewned distribution - expected but still nice to include

```{r}
shapiro.test(acs_test$FamilyIncome)
```

# Task 2: Preliminary EDA and feature engineering

Before trying to build any classification models, you should do some exploratory data analysis to
try to get a sense of which variables might be useful for trying to predict cases where FamilyIncome >= 250000. You should use a combination
of group by analysis (i.e. plyr or dplyr or similar) and plotting.  

If you decide you'd like to create some new variables (feature engineering), feel free to do that. Just document what you did and why you did it. 

```{r}
# Get some summary stats on each variable
summary(acs_fit)
```

# Histogram

```{r}
# see that those that those that own home correlate with higher incomes overall
ggplot(acs_fit) + geom_histogram(aes(x=own_home), fill = "gray")
```

# Scatterplots

```{r}
# scatter number of workers and family income
ggplot(data=acs_fit) + geom_point(aes(x=NumWorkers, y=FamilyIncome))

# scatter plot shows that those not on foodstamps tend to have higher income = duh, but relevant for model later
ggplot(data=acs_fit) + geom_point(aes(x=foodstamp_binary, y=FamilyIncome))

# plot shows that homes with 2 workers correlate with higher incomes vs other number of workers
ggplot(data=acs_fit) + geom_point(aes(x=NumWorkers, y=FamilyIncome))

# notice that there are very few observations with male head type. Female head has lower income and married highest incomes
ggplot(data=acs_fit) + geom_point(aes(x=family_type_cat, y=FamilyIncome))

# scatter house costs and family income - see that higher house costs correlate to higher incomes (slightly) - nothin major though
ggplot(data=acs_fit) + geom_point(aes(x=HouseCosts, y=FamilyIncome))
```

```{r}
# create matrix of scatterplots
# pairs(acs[,1:19])
```
# Boxplot

#### coor_cartesian -> Setting limits on the coordinate system will zoom the plot (like you're looking at it with a magnifying glass), and will not change the underlying data like setting limits on a scale will.
```{r}
# See that outliers begin roughly around income of $100,000
ggplot(data=acs_fit) + geom_boxplot(aes(x=NumWorkers, y=FamilyIncome))  + coord_cartesian(ylim = c(0, 350000))
```

# Density Plots

#### These show the density by variable on axis. These are useful to see the concentration range of values
```{r}
ggplot(acs_fit) + geom_density(aes(x=acs_fit$FamilyIncome)) + scale_x_continuous(labels=dollar)
```

```{r}
ggplot(acs_fit) + geom_density(aes(x=acs_fit$HouseCosts)) + scale_x_continuous(labels=dollar)
```
```{r}
ggplot(acs_fit) + geom_density(aes(x=acs_fit$NumChildren)) + scale_x_continuous()
```
```{r}
ggplot(acs_fit) + geom_density(aes(x=acs_fit$FamilyIncome)) + scale_x_log10(breaks =c(100,1000,10000,100000), labels=dollar) + annotation_logticks(sides="bt")
```
```{r}
ggplot(acs_fit) + geom_density(aes(x=acs_fit$HouseCosts)) + scale_x_log10(breaks =c(100,1000,10000,100000), labels=dollar) + annotation_logticks(sides="bt")
```
# Misc Plots
```{r}
# shows positive correlation between insurance and family income
ggplot(acs_fit, aes(x=acs_fit$Insurance, y=acs_fit$FamilyIncome)) +geom_point() + geom_smooth()
```
```{r}
# density plot for electrical bil
ggplot(acs_fit) + geom_density(aes(x=acs_fit$ElectricBill)) + scale_x_log10(breaks =c(100,1000,10000,100000), labels=dollar) + annotation_logticks(sides="bt")

# shows positive correlation between electric bill and family income
ggplot(acs_fit, aes(x=acs_fit$ElectricBill, y=acs_fit$FamilyIncome)) +geom_point() + geom_smooth()
```

# Group by analysis

```{r}
# This shows a good spread or range of each family type group. This will lend itself well to being included in my analysis
ddply(acs_fit,.(FamilyType),summarise,family_type_count=length(FamilyIncome))
```
```{r}
# Interesting look at mean income of family type grouped with home ownership type
ddply(acs_fit,.(FamilyType,OwnRent), summarise, mean_income=mean(FamilyIncome))
```
```{r}
ddply(acs_fit,.(FamilyType,FoodStamp), summarise, mean_income=mean(FamilyIncome))
```
```{r}
# simple look at mean income by foodstamp. Obvious results, but at the same time surprising to find that mean income 
# for those on food stamps in near $50k
ddply(acs_fit,.(FoodStamp), summarise, mean_income=mean(FamilyIncome))
```
```{r}
ddply(acs_fit,.(FoodStamp,NumBedrooms), summarise, mean_income=mean(FamilyIncome), num_bedrooms=mean(NumBedrooms))
```
```{r}
# This is a little excessive, but would be useful for piping it to a csv file (for example) if relevant to tasks at hand
ddply(acs,.(NumBedrooms,NumChildren,NumPeople,NumRooms,NumUnits,NumVehicles,NumWorkers), summarise, mean_income=mean(FamilyIncome))
```
```{r}
ddply(acs_fit,.(OwnRent), summarise, mean_income=mean(FamilyIncome))
```

# Count (family income) by various important indicators/variables

```{r}
# Family Type
tapply(acs_fit$FamilyIncome,acs_fit$FamilyType,length)
tapply(acs_fit$FamilyIncome,acs_fit$FamilyType,mean)
```

```{r}
# Own/Rent
tapply(acs_fit$FamilyIncome,acs_fit$OwnRent,length)
tapply(acs_fit$FamilyIncome,acs_fit$OwnRent,mean)
```

```{r}
# Insurance
tapply(acs_fit$FamilyIncome,acs_fit$FoodStamp,length)
tapply(acs_fit$FamilyIncome,acs_fit$FoodStamp,mean)
```